\chapter*{Declaration}
\thispagestyle{empty}
\noindent
%
I, \_\_\_\_\_\_\_\_\_\_ (Name), student registration number:\_\_\_\_\_\_\_\_, hereby confirm that I completed the submitted work independently and without the unauthorized assistance of third parties and without the use of undisclosed and, in particular, unauthorized aids. This work has not been previously submitted in its current form or in a similar form to any other examination authorities and has not been accepted as part of an examination by any other examination authority.
\\
Where the wording has been taken from other people’s work or ideas, this has been properly acknowledged and referenced. This also applies to drawings, sketches, diagrams and sources from the Internet.
\\
In particular, I am aware that the use of artificial intelligence is forbidden unless its use an aid has been expressly permitted by the examiner. This applies in particular to chatbots (especially ChatGPT) and such programs in general that can complete the tasks of the examination or parts thereof on my behalf.
\\
Furthermore, I am aware that working with others in one room or by means of social media represents the unauthorized assistance of third parties within the above meaning, if group work is not expressly permitted. Each exchange of information with others during the examination, with the exception of examiners and invigilators, about the structure or contents of the examination or any other information such as sources is not permitted. The same applies to attempts to do so.
\\
\\
Any infringements of the above rules constitute fraud or attempted fraud and shall lead to the examination being graded “fail” (“nicht bestanden”).
%
\vspace{2cm}
\\
\begin{minipage}[t]{0.45\textwidth}
    \rule{\textwidth}{0.5pt}\\
	Erlangen, date
\end{minipage}
\hfill
\begin{minipage}[t]{0.45\textwidth}
	\rule{\textwidth}{0.5pt}\\
	name\\
	91054 Erlangen\\
    address \\
\end{minipage}